\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-04-30 project-writings/paper.tex
Spell check.

\section{Introduction}
With the explosion of browser based applications that has occured in
recent years, Javascript has become one of the most widely used
programming languages, and Javascript performance has become
critically important. Unfortunately, Javascript seems to be designed
to be un-performant.

One source of these woes are Javascript's dynamic semantics for
numbers: all numeric values are defined by the specification to be
IEEE 754 double precision floating point values (with bitwise
operations being defined to round to the nearest 32 bit integer before
performing the operation). Since the common case is that numeric
values are integers and because floating point operations are
more expensive, all major Javascript implementations attempt to
represent numbers as integers when possible.

@@

The core of our analysis system is essentially an implementation of
Gough and Klaeren's static range analysis algorithm for bounds check
elimination.
We identify our technical contributions as:
\begin{itemize}
\item An implementation of Gough and Klaeren's range analysis in a
  modern industrial-strength SSA-based compiler.
\item A different algorithm for inserting and removing $\beta$ nodes
  that used by Gough and Klaeren that takes advantage of the data
  structures of a modern SSA compiler.
\item To facilitate the above, extended IonMonkey with a fast
  (constant time, one integer per basic block memory overhead) 
  mechanism for determining whether a block is dominated by another.
\item Extending the analysis to produce the information needed for
  overflow check elimination by differentiating between
  ``maximum/minimum 32-bit int values'' and ``infinity'' (possible
  overflow).
\item The observation that when doing range computation with the above
  distinction, the input ranges can be always be ``squashed'' to only
  finite ranges, since if overflow had occured, the code would have
  bailed out.
\item An implementation of overflow check elimination using the above
  analyses.
\end{itemize}

@@

The approach we took, again following \cite{Gough94eliminatingrange},
is to instead attack the problem in ``SSA style'', by introducing
$\beta$-nodes, a new form of pseudo-operation that associates range
information with a value. $\beta$ nodes take one argument and
additionall have an auxillary constant range associated with
them. Operationally, $\beta$-nodes are simply copies, but the
invariant is maintained that the output value will fall inside the
range given by the $\beta$ node. Gough and Klaeren refer to SSA
extended with $\beta$ nodes as XSA.

>>>>

\section{Introduction}
With the explosion of browser based applications that has occurred in
recent years, Javascript has become one of the most widely used
programming languages, and Javascript performance has become
critically important. Unfortunately, Javascript seems to be designed
to be un-performant.

One source of these woes are Javascript's dynamic semantics for
numbers: all numeric values are defined by the specification to be
IEEE 754 double precision floating point values (with bit-wise
operations being defined to round to the nearest 32 bit integer before
performing the operation). Since the common case is that numeric
values are integers and because floating point operations are
more expensive, all major Javascript implementations attempt to
represent numbers as integers when possible.

@@

The core of our analysis system is essentially an implementation of
Gough and Klaeren's static range analysis algorithm for bounds check
elimination.
We identify our technical contributions as:
\begin{itemize}
\item An implementation of Gough and Klaeren's range analysis in a
  modern industrial-strength SSA-based compiler.
\item A different algorithm for inserting and removing $\beta$ nodes
  that used by Gough and Klaeren that takes advantage of the data
  structures of a modern SSA compiler.
\item To facilitate the above, extended IonMonkey with a fast
  (constant time, one integer per basic block memory overhead) 
  mechanism for determining whether a block is dominated by another.
\item Extending the analysis to produce the information needed for
  overflow check elimination by differentiating between
  ``maximum/minimum 32-bit int values'' and ``infinity'' (possible
  overflow).
\item The observation that when doing range computation with the above
  distinction, the input ranges can be always be ``squashed'' to only
  finite ranges, since if overflow had occurred, the code would have
  bailed out.
\item An implementation of overflow check elimination using the above
  analyses.
\end{itemize}

@@

The approach we took, again following \cite{Gough94eliminatingrange},
is to instead attack the problem in ``SSA style'', by introducing
$\beta$-nodes, a new form of pseudo-operation that associates range
information with a value. $\beta$ nodes take one argument and
additionally have an auxiliary constant range associated with
them. Operationally, $\beta$-nodes are simply copies, but the
invariant is maintained that the output value will fall inside the
range given by the $\beta$ node. Gough and Klaeren refer to SSA
extended with $\beta$ nodes as XSA.
